---
title: "Speed VS Precipitation check"
author: "Mustafa Attallah"
date: "4/19/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)
```

```{r}
data<- read.csv("https://mathstat.slu.edu/~clair/stat/data/sensor_D.csv")
```

# visualizing the data 
```{r}
data %>%  ggplot(aes(x = precipitation, y = speed_km))+geom_point()
```
We can hardly see any pattern! big chunk of the data is on **zero** precipitation and that is because most of the time it is not raining and the speed drop might be because of something else, like rush hours and incidents. 


# Checking the normality of the Speed:
```{r}
hist(data$precipitation)
qqnorm(data$precipitation)
```



```{r}
#plot(density(data$speed_km), col = "green", main= "Density plot for Speed at D", ylim = c(0,0.1))
#data %>% filter(precipitation == 0) %>% with(lines(density(speed_km), col = "red"))
#data %>% filter(precipitation >10) %>% with(lines(density(speed_km), col ="blue"))
data %>% ggplot(aes(x= speed_km))+geom_density(color ="black")+geom_density(aes(color = (precipitation >0)))

# what do you think is gonna happen if we increase the precipitation filter values?
```

Hmm, see the **blue** line which is related to heavy rain observations! the speed isn't really distributed well here. 
We can see that filtering the data for **zero** precipitation didn't really change how the data look like! however, we see that the plot looks thinner! as the speed is relatively try to be higher. Thus, we cannot apply t test on such skewed data. wilcoxon is a good approach to think about!



### the hypotheses
$H_o: \mu = 96.56$
$H_a: \mu \ne 96.56$

```{r}
with(data, wilcox.test(speed_km, conf.int = T,mu = 96.56))

data %>% 
  filter(precipitation > 10) %>% 
  with(wilcox.test(speed_km, conf.int = T,mu = 96.56))


```

**Decision** : There is enough evidence to **reject** the null hypothesis as **p_value** is less than **0.05**. 

Interpretation: according to wilcoxen test above, We believe that the average speed is less than 96.56 km/h when it is raining. 




In this step, a model is being built to check if there is any kind of correlation between the speed and precipitation rate.

```{r}
mod_D<- lm(speed_km~precipitation, data=data)
summary(mod_D)
ggplot(data, aes(x = precipitation, y = speed_km))+geom_point()+geom_smooth(method = "lm")
```
Some interpretations:
```{r}
plot(mod_D)
data[520,]
```


