---
title: 'Homework #02'
author: "Robert Campbell"
date: "16 May 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
suppressWarnings(suppressMessages(library(dplyr)))
# Chapter 2 exercises # 1-5, 8-11, 17, 20, 21, 25, 26, 29
```

## Chapter 02   
### Problem 01      
  6/36 or 1/6
```{r}
mean(replicate(10000, {x<-sample(1:6, 2, replace=TRUE)
  ((x[1] == (2 * x[2])) || (x[2] == (2* x[1])))
  }))
```
   
### Problem 02      
  a. 6/36 or 1/6
```{r}
mean(replicate(10000, {x<-sample(1:6, 2, replace=TRUE)
  abs(x[1] - x[2]) == 0
}))
```   
  b. 4/36 or 1/9
```{r}
mean(replicate(10000, {x<-sample(1:6, 2, replace=TRUE)
abs(x[1] - x[2]) == 4
}))
```   
   
### Problem 03      
  a. P(not green) = 1 - .2 = .8
  b. P(r U o U y) = .13 + .2 + .14 = .47
  c. Approximately .2
```{r}
mm<- c("y","r","o","br","g","bl")
pb<-c(.14,.13,.2,.12,.2,.21)
mean(replicate(10000, {x<-sample(mm, 4, prob=pb, replace=TRUE);
  sum(x=="bl") > 1
}))
```   
  d. Approximately .01
```{r}
mean(replicate(10000, {x<-sample(mm, 6, prob=pb, replace=TRUE);
  ((sum(x=="y") == 1) & (sum(x=="r") == 1) & (sum(x=="o") == 1) &
    (sum(x=="br") == 1) & (sum(x=="g") == 1) & (sum(x=="bl") == 1)) 
}) )
```


### Problem 04      
  Approximatly .06   
```{r}
mean(replicate(10000, {x<-sample(mm, 30, prob=pb, replace=TRUE);
  ( (sum(x=="bl") > 8) & (sum(x=="o") > 5)) 
}) )
```   


### Problem 05      
Approximately 0.37 or 3/8   
```{r}
type<- c("o","a","b","ab")
pb<-c(.45,.40,.11,.04)

mean(replicate(10000, {x<-sample(type,2,prob=pb,replace=TRUE);
  x[1] == x[2]
}))
```
   

### Problem 08      
  Approximately 0.56
```{r}
mean(replicate(10000, {x<-sample(1:6, 5, replace=TRUE)
  ((14 < sum(x)) & (sum(x) < 21))
}))
```   


### Problem 09      
  Approximately 0.29
```{r}
mean(replicate(10000, {x<-sample(1:6, 20, replace=TRUE)
  y<-cumsum(x)
  sum(y == 20)
}))
```   


### Problem 10      
  a. 
```{r}
y <- replicate(10000, {x<-sample(1:6, 2, replace=TRUE)
sum(x)
})

hist(y, breaks=12)
```   
  b. 
```{r}
y <- replicate(10000, {x<-sample(c(2,6), replace=TRUE)
  x + 5
})

hist(y)
```   
  c. 
```{r}
y <- replicate(10000, {x<-sample(c(1,2,2,3,3,4), replace=TRUE)
  z<-sample(c(1,3,4,5,6,8), replace=TRUE)
  x + z
})

hist(y, breaks=12)

```   

### Problem 11      
  Assuming birth year is not accounted for, excluding Feb29
  Approximately 0.10
```{r}
mean(replicate(10000, {x<-sample(1:365, 200, replace=TRUE)
  sum(x==1) > 1
}))
```   
   

### Problem 17      
  Approximately 0.02
```{r}
s<-fosdata::scrabble

mean(replicate(10000, {x<-sample(s$piece,7,replace=FALSE)
  ((sum(x=='A')==0)&(sum(x=='E')==0)&(sum(x=='I')==0)&(sum(x=='O')==0)&(sum(x=='U')==0))
}))
```   


### Problem 20      
  a. C and D are disjoint, A and D are disjoint
  b. B and D are independent
  c. A and B, A and C, B and C


### Problem 21      
  a. A and C
  b. B and C
  c. 1


### Problem 25      
  a. 1/2
  b. p^2 * (1-p)^2 = p^2 * (1 - 2p + p^2) = p^3 - 2p^2 + p
  c.
```{r}
p<-seq(0,1,0.1)
y<- p^3 - 2*(p^2) + p
plot(p, y)
```      
  d. Prob same is approximately .51, different is approximately .49
```{r}
flip<-c("v", "c")
pb<-c(.45,(1-.45))

mean(replicate(10000, {x<-sample(flip,2, prob=pb,replace=TRUE)
  (x[1]==x[2])
}))
```   
  e. Choose same
  f. Yes, the two frisbee flip is closer to a .5 probability   


### Problem 26      
  Approximate .5
```{r}
people <- sample(0:1, 100000, prob=c(.99,.01), replace=TRUE)
test <- ifelse(people, 1, sample(0:1, 100000, prob=c(.99,.01), replace=TRUE))
1-(sum(test & !people)/sum(test))
```   


### Problem 29      
  a.   
```{r}
# P(B|A) = P(A|B) * (P(B)/P(A))
# If there is a second tree, there will always be a first tree
  1 * (.85/.91)
```   
  b. 



