---
title: 'Homework #08'
author: "Robert Campbell"
date: "12 Apr 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
library(dplyr)
library(ggplot2)
#library(babynames)
library(tidyr)
#library(stringr)
library(ISwR)
library(Sleuth3)

# Homework Problems:
# Ch 8 Exercises # 14, 16, 17, 19, 25-32, 34, 35, 21, 23, 24
```


## Chapter 08
### Problem 14   

98% Confidence interval: [-1.037, -0.556]
P value is < 0.05, so the mean differs significantly from 0. This means that the two nurses' interpretations of the data differed significantly, effecting the values.
```{r}
react <- ISwR::react
xbar <- mean(react)
s <- sd(react)
stderror <- s/sqrt(length(react))
tcritical <- -qt(0.01,(length(react)-1))
xbar - tcritical * stderror  # lower
xbar + tcritical * stderror  # upper
t.test(react,conf.level = 0.98)
```

### Problem 16   
Hypothesis: Children's estimate of an object's weight will deviate from 0, meaning they will make inaccurate estimates.

The only accurate estimates were for the 200g object, all others were significantly different.
```{r}
child_est<- fosdata::weight_estimate %>% filter(age!= 'adult')
t.test(child_est$mean100, mu=100)
t.test(child_est$mean200, mu=200)
t.test(child_est$mean300, mu=300)
t.test(child_est$mean400, mu=400)
```

### Problem 17    
Population is Mexican-American adults in California, with a sample size of 102.
P-value = (1.743 x 10^-4) < 0.05, so the blood pressure significantly differs from the normal systolic blood pressure of 120 mm Hg.
```{r}
obesity<-ISwR::bp.obese
t.test(obesity$bp, mu=120)
```

### Problem 19
Natural null hypothesis is that the population's ratio of actual weight to ideal weight will not deviate from the mean of 1.

P-value is much less that 0.05, meaning that the population's mean ratio significantly differs from the null hypothesis.
```{r}
obesity<-ISwR::bp.obese
t.test(obesity$obese, mu=1)
```

### Problem 25   
  a. Pain felt by a laser shot on the hand will not significantly change based on the aesthetic quality of an observed painting.   
  b. There is no difference in appearance of objects between normal viewing and viewing by bending over and looking between your legs.   
  c. The time to urinate will differ based off of species.

### Problem 26   
  a. The mean of the sample is greater than the null hypothesis mean in all measurements.
  b. The quality of sleep did not see significant improvement. This can be seen by the P-value being greater than 0.05 (P = 0.27).   
  c. Reported sleep disturbance showed the most significance, having the smallest P-value of less than 0.01.

### Problem 27   
  P-value = 0.6463 > 0.05, there is no significant difference from between male and female blood pressure in the sample group.   
  It would not be appropriate to generalize this to the general American population. It is a single ethnic group, in a single geographic region, in a "small town", and does not account for the diversity in each of those larger categories that make up the general American population.

```{r}
obesity<-ISwR::bp.obese
t.test(bp ~ sex, data=obesity)
```

### Problem 28    
  There is significant evidence, with the p-value = 0.01308. With Fish Oil's mean reduction in diastolic blood pressure having a positive value while regular oil has a negative value showing that fish oil provides greater reduction in diastolic blood pressure on average. 
```{r}
oil<-Sleuth3::ex0112
t.test(BP ~ Diet, data=oil)
```

### Problem 29  
  No, the initial boxplot looks lopsided and the initial histograms look more exponential than normal.   
  After calculating the logs, the data does appear more normal.   
  The difference in mean of the log of brainsize between large and small litter size is not significant as P-value= 0.05225 is not less than 0.05, though it is close.

```{r}
size <- Sleuth3::ex0333
boxplot(BrainSize ~ LitterSize, data=size)
size %>% ggplot(aes(x=BrainSize)) + geom_histogram() + facet_grid(vars(LitterSize))
size<-mutate(size,logBrain=log(BrainSize))
boxplot(logBrain ~ LitterSize, data=size)
size %>% ggplot(aes(x=logBrain)) + geom_histogram() + facet_grid(vars(LitterSize))
t.test(logBrain ~ LitterSize,data=size)
```

### Problem 30   
  Hypothesis: There is a significant difference in volume in the affected brain regions of schizophrenic patients.
  The sample data is inconclusive with P-Value = 0.05691 being greater than 0.05, but close.
  
```{r, warning=FALSE, message=FALSE}
vol<-Sleuth3::case0202
pvol<-vol %>% pivot_longer(col=Unaffected:Affected, names_to = "Diagnosis", values_to = "BrainVolume")
pvol %>% ggplot(aes(x=BrainVolume)) + geom_histogram() + facet_grid(vars(Diagnosis))
t.test(BrainVolume ~ Diagnosis, pvol)
```

### Problem 31   
  a. The tk-12 provides a significantly different temperature change from the control with a P-value of (5.609 x 10^-15), and provides a cooling afffect as all data values are negative while the control are positive.
```{r}
moo<-fosdata::cows_small
t.test(moo$tk_12, moo$control, paired = TRUE)
```

  b. The tk-0.75 provides a significantly different temperature change from the control with a P-value of (3.747 x 10^-14), and provides a cooling effect as all data values are negative while the control are positive.
```{r}
t.test(moo$tk_0_75, moo$control, paired = TRUE)
```


### Problem 32   
  There is not a significant difference in task completion time based on gender as the P-value: 0.2033 much larger than the alpha: 0.01

```{r}
child<-fosdata::child_tasks
t.test(day_night_completion_time_secs ~ gender, data=child)
```

### Problem 21   
  a. The mean of 4 was within the confidence interval. (May not show that way after the knit)   
```{r}
x<-rnorm(10,mean=4, sd=1)
t.test(x,conf.level = 0.95)
```
  b. 
```{r}
z<-replicate(10000, {x<-rnorm(10,mean=4, sd=1);
  y<-t.test(x,conf.level = 0.95)$conf.int;
  (4 > y[1]) && (4<y[2])})
mean(z)
```

### Problem 23   
  a. The mean of 4 did appear in the confidence interval. (This may not match the R output after knitting)
```{r}
x<-rexp(10,0.25)
t.test(x,conf.level = 0.95)
```
  b. The result is not .95 because the data is heavily skewed left.
```{r}
z<-replicate(10000, {x<-rexp(10,0.25);
  y<-t.test(x,conf.level = 0.95)$conf.int;
  (4 > y[1]) && (4<y[2])})
mean(z)
```
  c. This is closer to .95 because of the larger sample in each experiment iteration, which brings the sample mean closer to the true mean.
```{r}
z<-replicate(10000, {x<-rexp(100,0.25);
  y<-t.test(x,conf.level = 0.95)$conf.int;
  (4 > y[1]) && (4<y[2])})
mean(z)
```

### Problem 24   
```{r}
pval<- replicate(10000,{x<-rnorm(20,0,1);
  y<-convolve(x, c(1, 1), type = "filter");
  t.test(y, mu=0)$p.value})
mean(pval>0.05)
```


