---
title: "Exam 1"
author: "Robert Campbell"
date: "Friday, March 5, 2021"
output: pdf_document
---

Place your answers into this markdown document, knit it, and hand in the result as a PDF.
There are 9 questions, each worth 10 points.  Additionally, you get 10 points for handing in a PDF of a knit markdown document.

You may use R, the internet, and any reference material,
but do not work together and do not get help (except from Dr. Clair).

### Honor Pledge
The work I have submitted represents my own effort. While working on this exam, I did not communicate in any form with individuals other than the instructor.

Signed: Robert V. Campbell

### Problem 1
Suppose you roll an ordinary six sided die 20 times and add up the total.
What is the probability the sum is 55 or less?
```{r}
x<-replicate(10000,sum(sample(1:6,20,replace=TRUE)))
mean(x<56)
```


### Problem 2
McDonald's happy meals currently contain a package of four Pokemon cards. There are 25 different cards, all equally likely.

a. If you buy one happy meal (4 cards), what is the probability you get a Pikachu card?
```{r}
x<-replicate(10000,sum(sample(1:25, 4, replace=TRUE) == 1))
mean(x>0)
```

b. If you buy 10 happy meals, what is the probability you get a Pikachu card?
```{r}
x<-replicate(10000,sum(sample(1:25, 40, replace=TRUE) == 1))
mean(x>0)
```
c. How many happy meals do you need to buy to have a 99% chance of getting a Pikachu?
```{r}
x<-replicate(10000,sum(sample(1:25, 116, replace=TRUE) == 1))
mean(x>0)
116/4
```


### Problem 3
Roll two dice and consider the events

- $A$: both dice are even
- $B$: the sum of the dice is 8

a. What is $P(A)$?
  $P(A) = \frac{9}{36} = \frac{1}{4}$
b. What is $P(A|B)$?
   $P(A|B) = \frac{P(A\&B)}{P(B)} = \frac{3/36}{5/36}=\frac{3}{5}$
c. Are $A$ and $B$ independent events?
  No.

### Problem 4

The random variable $X$ has the pdf shown here:

```{r echo=FALSE}
px <- seq(-1,3,.1)
plot(px,pmax(1-abs(px-1),0),type='l',xlab="X",ylab="")
```

a. What is $P(X < 1)$?
  $P(X < 1) = \frac{1}{2}$
b. What is $P(X > 1.5)$?
  $P(X > 1.5) = \frac{1}{8}$
c. What is $E[X]$?
  $E[X] = 1$

### Problem 5

A high voltage spark can be used to create tiny aerosol particles of gold.
These particles have diameters which vary according to a normal distribution with mean 1.35nm and standard deviation 0.24nm.

a. What proportion of these particles have diameter less than 1nm?
```{r}
x<-rnorm(10000,1.35,0.24)
mean(x<1)
```

b. What proportion of these particles have diameter larger than 2nm?
```{r}
mean(x>2)
```

### Problem 6
Back in 2006, the website MySpace leaked a large number of its user's passwords.
The passwords had this distribution:

- Numbers only: 1.3%
- Letters only: 9.5%
- Alphanumeric (both letters and numbers): 81%
- Includes symbols: 8.2%

a. What is the probability that a MySpace user had no Symbols in their password?
  $1-.082=  0.918$

b. If you chose 10 MySpace users at random, what is the probability that all of them had Alphanumeric passwords?

```{r}
x<- replicate(10000,sum(sample(c('N','L','AN','S'), 10, replace=TRUE,prob=c(.013,.095,.81,.082)) == 'AN'))
mean(x==10)
```
### Problem 7

Let $X$ and $Y$ be normal random variables with mean 10 and sd 5.

Find the expected distance from the point $(X,Y)$ to the origin $(0,0)$.
```{r}
sqrt((10^2)+(10^2))

```

### Problem 8

The `ChestSizes` data from the `HistData` library has measurements of Scottish Militiamen.
There are two variables: `chest`, which gives the chest size in inches, and `count` which gives the number of soldiers with that chest size.

a. Calculate the total number of soldiers by summing the `count` variable.
```{r}
sum(HistData::ChestSizes$count)

```

b. Calculate the mean chest size by summing (chest size)*(count)/(total number of soldiers).
```{r}
sum((HistData::ChestSizes$count)*(HistData::ChestSizes$chest))/sum(HistData::ChestSizes$count)
```

### Problem 9
Continue with `HistData::ChestSizes`.

a. Plot the distribution of chest sizes with `chest` on the $x$-axis and `count` on the $y$-axis.
```{r}
plot(HistData::ChestSizes$chest,HistData::ChestSizes$count)
```
b. What sort of random variable could model this distribution?
  Normal RV

c. From your plot, estimate (by eye) the mean and standard deviation of chest sizes.
  Mean: 40, sd: 2