---
title: 'Homework #03'
author: "Robert Campbell"
date: "22 Feb 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
suppressWarnings(suppressMessages(library(dplyr)))
# Homework Problems:
# 1, 3, 5, 7, 9, 11, 14, 15, 16, 17, 19, 21, 23

```

## Chapter 03

### Problem 01
  a. p(x) is a valid probability mass function 
     $\frac{1}{4} + \frac{1}{2} + \frac{1}{8} + \frac{1}{8} = 1$
  b. $P(X\geq2) = \frac{1}{8} + \frac{1}{8} = \frac{1}{4}$
  c. $P(X\geq2 | X\geq1) =\frac{1}{2} + \frac{1}{8} + \frac{1}{8} = \frac{5}{8}$
  d. $P(X\geq2 \bigcup X\geq1) =\frac{1}{2} + \frac{1}{8} + \frac{1}{8} = \frac{5}{8}$
  
### Problem 03
  $1 = C(1+\frac{1}{2}+\frac{1}{4})$
  $\frac{1}{C} = \frac{7}{4}$
  $C = \frac{4}{7}$

### Problem 05
```{r}
X <- c(0,1,2,3)
prob <- c(1/4, 1/2, 1/8, 1/8)
mean(sample(X, 10000, prob, replace=TRUE))
```

### Problem 07
```{r}
d1 <- sample(c(1,2,3,4,5,6), 10000, replace=TRUE)
d2 <- sample(c(1,2,3,4,5,6), 10000, replace=TRUE)
mean(d1*d2)
```

### Problem 09
  0:999 has 1000 numbers with 10 possible 2 digit numbers(0:9), 90 possible 3 digit numbers(10:99), and 900 3 digit numbers(100:999).
  ```{r}
  mean(sample(c(1,2,3), 10000, prob=c(1/1000, 90/1000, 900/1000), replace=TRUE))
  ```

### Problem 11
  Bernoulli trial, red being a win, with p = 18/38
  a. Expected value is np, so the expected payout would be \[\$1 * \frac{18}{38} = \$0.47\]
  b. $P(A|B)=\frac{P(A\&B)}{P(B)}= \frac{\frac{18}{38}*\frac{18}{38}}{\frac{18}{38}}=\frac{18}{38}$
     The expected payout would be $\$2 * \frac{18}{38} = \$0.97$

### Problem 14
Note: According to the website's probabilities, you can win a front pair and a straight so I am treating the "don't care" portions of the probabilities ignoring the better results. The other option would be to change front and back pair to 9/1000 instead of 1/100. This same logic will carry throughout the problem.
  a. $\frac{1}{10} * \frac{1}{10} * \frac{1}{10} = \frac{1}{1000}$
  b. $\frac{1}{10} * \frac{1}{10} * \frac{10}{10} = \frac{1}{100}$
  c. $\frac{10}{10} * \frac{1}{10} * \frac{1}{10} = \frac{1}{100}$
  d. $\frac{3}{10} * \frac{2}{10} * \frac{1}{10} = \frac{6}{1000}$
  e. 
  f. 

### Problem 15
$E[X] = \sum_1^k(X*p(x)) = \frac{k(k+1)}{2}$

### Problem 16
  a. 
  ```{r}
  test <- rbinom(10000, 20, .25)
  (mean(test)/20)*100
  ```
  b. 
  ```{r}
  (sum(test>=10)/10000)
  ```
### Problem 17
  a. 
  ```{r}
  x <- rbinom(10000,10, .91)
  mean(x)
  ```
  b.
  ```{r}
  (sum(x >= 8)/10000)
  ```
  
### Problem 19
X <- rbinom(trials, 200, .55)
plot(table(X))
mean(X)
var(X)
# sd is sqrt(np(1-p))
sd(X)
# margin of error
2*sd(X)/200
X < 100 # This is the event that prop A looks like it will fail

  a. Expected number of votes:
  ```{r}
  x <- rbinom(10000, 200, .55)
  mean(x)
  ```
  b. Margin of error in percent of votes:
  ```{r}
  (2*sd(x))/200 * 100
  ```
  c. Probability prop A will fail?
  ```{r}
  (sum(x<100)/10000)
  ```
  d.
  $2\% = \frac{2*\sqrt{n(.55)(1-.55)}}{n}$
  $(.01n)^2 = n(.55)(.45)$
  $n = $

### Problem 21
  a. $P(y): 1/4, y = -1; 1/2, y = 0; 1/8, y =1; 1/8, y=2$
  b. $P(u): 1/4, u = 0; 1/2, u = 1; 1/8, u=4; 1/8, u = 9$
  c. $P(v): 1/2, v=0; 3/8, v = 1; 1/8, v = 4$

### Problem 23
```{r}
X <- c(0,1,2,3)
prob <- c(1/4, 1/2, 1/8, 1/8)
y <- sample(X, 10000, prob, replace=TRUE)
```
  Variance:
  ```{r}
  var(y)
  ```
  Standard Deviation:
  ```{r}
  sd(y)
  ```