---
title: 'Homework #07'
author: "Robert Campbell"
date: "31 Mar 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
library(dplyr)
#library(ggplot2)
#library(babynames)
#library(tidyr)
#library(stringr)

# Homework Problems:
# Ch 5 Exercises # 25, 26, 28, 30, 34
# Ch 8 Exercises # 1, 3, 4, 6, 7, 11, 13
```


## Chapter 05
### Problem 25  
N = 8
```{r}
n<-8
trials<-10000
mu<-.5*(0+1)
sigma<- sqrt((1/12)*(1-0)^2)
sim<-replicate(trials, {x<-runif(n,0,1);
(mean(x)-mu) / (sigma/sqrt(n))
})
plot(density(sim))
curve(dnorm(x),add=TRUE,col="red")
```

### Problem 26  
By inspection, mu = 10, sigma = 2.5
```{r}
mu<-10
sigma<-2.5
n<-100
sumNorm<-replicate(10000,{x <- replicate(n, x<-sum(rexp(20, 2)));
                  (mean(x)-mu)/(sigma/sqrt(n))})
plot(density(sumNorm))
```

### Problem 28  
N = 36
```{r}
n<-36
trials<-10000
k<-2
mu<-k
sigma<- sqrt(2*k)
sim<-replicate(trials, {x<-rchisq(n,df=k);
(mean(x)-mu) / (sd(x)/sqrt(n))
})
plot(density(sim))
curve(dnorm(x),add=TRUE,col="red")
```

### Problem 30  
#### a.  

```{r}
curve(dt(x, 1), from = -5, to = 5)
```

#### b.  
```{r}
x<-replicate(10,mean(rt(100,1)))
y<-replicate(10,mean(rt(1000,1)))
z<-replicate(10,mean(rt(10000,1)))
x;y;z
```

#### c.  

```{r}
plot(density(x))
plot(density(y))
plot(density(z))
```

### Problem 34  
```{r}
n<-8
mu<-2
sim<-replicate(10000, {x<-rnorm(n,2,3);
(mean(x)-mu) / (sd(x)/sqrt(n))
})
plot(density(sim))
curve(dt(x,df=7),add=TRUE,col="red")
```


## Chapter 08
### Problem 01  
```{r}
n<-12
mu<-1
sigma<-3
sim<-replicate(10000, {x<-rnorm(n,1,sigma);
(mean(x)-mu) / (sigma/sqrt(n))
})
plot(density(sim))
curve(dt(x,df=11),add=TRUE,col="red")
```

### Problem 03  
C = 1.439756
```{r}
qt(0.1, df=6)
```

### Problem 04  
The smaller the degrees of freedom, the more area in the tail, the greater the variance.
```{r}
1-pnorm(2)
1-pt(2,df=40)
1-pt(2,df=20)
1-pt(2,df=10)
1-pt(2,df=5)

curve(dnorm(x), from = -5, to = 5, lwd=2)
curve(dt(x,df=40),add=TRUE,col="red")
curve(dt(x,df=20),add=TRUE,col="blue")
curve(dt(x,df=10),add=TRUE,col="green")
curve(dt(x,df=5),add=TRUE,col="yellow")
```

### Problem 06  
99% confidence interval [16.49394,20.0219]
```{r}
plastics <- fosdata::plastics %>% filter(!is.na(diameter))
diam <- plastics$diameter
xbar<-mean(diam)
s<-sd(diam)
stderror<- s/sqrt(length(diam))
tcritical = -qt(0.005, df=(length(diam)-1))
xbar - tcritical * stderror
xbar + tcritical * stderror
```

### Problem 07
95% Confidence Interval: [299836.7, 299868.1]  
This interval does not include 299729m/s
```{r}
light <- morley$Speed + 299000
xbar<-mean(light)
s<-sd(light)
stderror<- s/sqrt(length(light))
tcritical = -qt(0.025, df=(length(light)-1))
xbar - tcritical * stderror
xbar + tcritical * stderror
```

### Problem 11  
a -> II.  
b -> I.  
c -> III.

### Problem 13  
#### a.
Population mean is 2
```{r}
lambda<-0.5
mu <- (1/lambda)
mu
```
#### b.
```{r}
z<-replicate(10000, {y<-t.test(rexp(10,lambda),mu=2)$conf.int;
      (mu >= y[1]) & (mu <=y[2])})
sum(z)/length(z)

```
#### c.
The value is different than 95% because it is pulling a very small sample, which can lead to very skewed means. Upping the sample size (not the replicate trials) increases the accuracy, while the larger number of trials increases precision.  
By reducing the replicate trials by a factor of ten and increasing the sample size by the same, we start seeing more accurate but less precise results.

```{r}
z<-replicate(1000, {y<-t.test(rexp(100,lambda),mu=2)$conf.int;
      (mu >= y[1]) & (mu <=y[2])})
sum(z)/length(z)
```