---
title: 'Homework #06'
author: "Robert Campbell"
date: "24 Mar 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(babynames)))
suppressWarnings(suppressMessages(library(tidyr)))
suppressWarnings(suppressMessages(library(stringr)))
# Homework Problems:
# Ch 6 Exercises # 1,4,5,8,9,13,15,19-23,26,28
```


## Chapter 06
### Problem 01
```{r}
quakes %>% ggplot(aes(x=lat, y=long, color=depth, size=mag)) + geom_point()
```

### Problem 04
```{r}
fosdata::brake %>% ggplot(aes(x=p1_p2)) + geom_density()
```
  a. The data is skewed
  b. The most likely time appears to be about 1125ms
  
### Problem 05
```{r}
fosdata::brake %>% ggplot(aes(sample=latency_p1)) + geom_qq() + geom_qq_line()
```
The data appears to be normal, but right skewed

### Problem 08
```{r}
fosdata::austen %>% filter(novel=="Pride and Prejudice") %>% group_by(chapter) %>% 
  summarize(sentiment = mean(sentiment_score)) %>%
  ggplot(aes(x=sentiment, y=chapter, color=chapter)) + geom_col()
```

### Problem 09
```{r}
suppressWarnings(suppressMessages(ISwR::juul %>% ggplot(aes(x=age, y=igf1,color=tanner)) +
                   geom_point() + geom_line() + geom_smooth()))
```
  c. Smooth seems to fit as the data does seem to more or less follow the trendline
     Line does not seem appropriate as the data jumps around that trendline in clusters

### Problem 13
```{r}
suppressWarnings(suppressMessages(fosdata::pres_election %>% mutate(party=factor(party)) %>%
  group_by(party, year) %>% summarize(votes=sum(candidatevotes,na.rm=TRUE)) %>%
  ggplot(aes(x=party, y=votes,fill=party)) + geom_col() + facet_wrap(vars(year))))
```

### Problem 15
  a.
```{r}
Lahman::Batting %>% group_by(yearID) %>% ggplot(aes(x=yearID, y=X2B, color=yearID)) + geom_point()
```
  b.
```{r}
suppressWarnings(suppressMessages(Lahman::Batting %>% group_by(yearID,lgID) %>%
  filter(lgID == c("AL", "NL")) %>% ggplot(aes(x=yearID, y=X2B, color=lgID)) +
    geom_point() + scale_color_manual(values = c("red1","blue"))))
```

### Problem 19
```{r}
suppressWarnings(suppressMessages(babynames %>% group_by(year, sex) %>% summarize(births = sum(n)) %>% 
  ggplot(aes(x=year, y=births, color=sex)) + geom_line()))
```
### Problem 20
```{r}
suppressWarnings(suppressMessages(babynames %>% group_by(year, sex) %>%
  summarize(u_names = n()) %>% ggplot(aes(x=year, y=u_names, color=sex)) +
    geom_line()))
```

### Problem 21
```{r}
babynames %>% filter(name %in% "Robert", sex=="M") %>%
  ggplot(aes(x=year,y=n)) + geom_line()
```

### Problem 22
```{r}
suppressWarnings(suppressMessages(babynames %>% filter(name %in% c("Bryan", "Brian"), sex=="M") %>%
  ggplot(aes(x=year,y=n, color=name)) + geom_line()))
```

### Problem 23
```{r}
suppressWarnings(suppressMessages(babynames %>% filter(name %in% "Jessie") %>%
  ggplot(aes(x=year,y=n, color=sex)) + geom_line()))
```
  - More male than female (~1950-1980)
  - More female than male (prior to 1950)
  - About the same (1980 and later)
  
  
### Problem 26
  a.
```{r, echo=FALSE}
tidy_data<-fosdata::scotland_births %>% pivot_longer(!age, names_to="Year", values_to="Births")
tidy_data$Year<-str_remove(tidy_data$Year, "x")
tidy_data$Year<-as.integer(tidy_data$Year)
tidy_data %>% group_by(age,Year) %>% ggplot(aes(x=Year,y=Births, group=age)) + geom_line()
```
  b.
```{r}
tidy_20<-tidy_data %>% filter(age == 20)
tidy_30<-tidy_data %>% filter(age==30)
tidy_data %>% group_by(age,Year) %>% ggplot(aes(x=Year,y=Births, group=age)) +
  geom_line() +geom_line(data=tidy_20, aes(x=Year,y=Births),color='red')+
geom_line(data=tidy_30, aes(x=Year,y=Births),color='blue') + labs(title="Scotland Births Over Time by Mother's Age \nRed: Age 20 \tBlue: Age 30")
```
### Problem 28

