---
title: 'Homework #11'
author: "Robert Campbell"
date: "10 May 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
library(dplyr)
library(ggplot2)
#library(babynames)
#library(tidyr)
#library(stringr)

# Homework Problems:
# Ch 12 Exercises # 4ab, 5, 6, 8, 9, 15
```


## Chapter 12   
### Problem 04   
  a. Looks like the data satisfies ANOVA assumptions
```{r}
pen<-palmerpenguins::penguins
boxplot(bill_depth_mm ~ species, data=pen)
```
   
  b. p-value < 2.2*(10^-16), showing significance at the alpha = 0.01 level.
      Reject the null, species does significantly predict bill depth.
```{r}
depth_mod<-lm(bill_depth_mm ~ species, data=pen)
anova(depth_mod)
```


### Problem 05  
  P-value: 0.00362, reject the null. There are significant differences
in the predictions of 300g objects by age group.
```{r}
anova(lm(mean300 ~ age, data=fosdata::weight_estimate))
```

### Problem 06  
  a. Spock's Lawyers may have a point, as Spock's mean percentage is visibly less than the means of the other 7 lawyers.     
```{r}
judge_perc <- Sleuth3::case0502
boxplot(Percent ~ Judge, data=judge_perc)
```
  b. P-value: 0.3239, do not reject the null. The means of the percentage of women jurors in the 6 other judges courts are significantly similar.   
```{r}
judge_perc %>% filter(Judge != "Spock's") %>% lm(Percent ~ Judge, data=.) %>% anova()
```
  c. The significantly similar means of the other judges justifies comparing Spock's judge to the other 6.
    P-value: 1.303 * (10^-6), reject the null. There is significant difference in the expected percentage of women jurors in Spock's judge's court when compared to the other judges.

```{r}
judge_perc %>% mutate(Spock = (Judge=="Spock's")) %>%
  t.test(Percent ~ Spock, data=.) 

```


### Problem 08  
  The p-value: 5.93*(10^-10) looks to show significant difference in weight based on feed, however the bloxplot shows inconsistent variance in the different groups with the most extreme being when you compare "meatmeal" with "sunflower".   

```{r}
anova(lm(weight ~ feed, data=chickwts))
boxplot(weight ~ feed, data=chickwts)
```

### Problem 09  
  With P-value: 0.0004827, it looks like there is significant difference in the mean based on experiment number. However, looking at the boxplots the data does not satisfy the anova assumptions, especially experiments 3 and 5.   
```{r}
anova(lm(Speed ~ Expt, data=morley))
boxplot(Speed ~ Expt, data=morley)
```

### Problem 15  
  a.   
```{r}
flint_pb <- fosdata::flint
boxplot(log(Pb1) ~ Ward, data=flint_pb)
flint_pb<- flint_pb %>% filter(Ward != 0)
boxplot(log(Pb1) ~ Ward, data=flint_pb)
```
  b. The log levels look normally distributed and the variances a similar enough to justify anova.   
  c. P-value: 0.03137, there is a significant difference in the mean of the log of the amount of lead in the water based on ward in Flint, MI. 
```{r}
anova(lm(log(Pb1) ~ Ward, data=flint_pb))
```
