---
title: 'Homework #10'
author: "Robert Campbell"
date: "15 May 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# This line makes your figures smaller (you want to do this!)
knitr::opts_chunk$set(fig.width=5, fig.height=3)

# This code loads the dplyr library
# You often want to load libraries at the start of your file
library(dplyr)
library(ggplot2)
library(quantreg)
#library(babynames)
#library(tidyr)
#library(stringr)

# Homework Problems:
# Ch 11 Exercises # 2-6, 10, 14-17, 20-25, 28
```


## Chapter 11   
### Problem 02    
  a. 
```{r}
faith_mod<-lm(waiting ~ eruptions, data=faithful)
faith_mod
```
  b.
```{r}
faithful %>%  ggplot(aes(x=eruptions, y=waiting)) + geom_point() +
  geom_smooth(method="lm")

```

  c.
```{r}
predict(faith_mod, newdata = data.frame(eruptions=4.3))
```  

### Problem 03    
  a.
```{r}
j <- ISwR::juul
j<-j %>% filter(tanner==5, age<20)
model <- lm(igf1 ~ age, data=j)
model
```
  b. igf1 = (-38.94)*age + 1135.49
  c. 
```{r}
predict(model, newdata=data.frame(age=16))
```

### Problem 04    
  a. Greater than .03
  b. Between  -.03 and .03
  c. Less than -.03   
  d. Greater than .03   


### Problem 05    
```{r}
d<-datasauRus::datasaurus_dozen
d %>% group_by(dataset) %>% summarize(correlation = cor(x, y, use="complete.obs"),
        xbar=mean(x), sx = (sd(x)/sqrt(142)), ybar=mean(y), sy=(sd(y)/sqrt(142)))
```
   
   They all seem to have pretty similar averages and stderr. X and Y seem to largely be not correlated.   

```{r}
d %>% ggplot(aes(x=x, y=y, color=dataset)) + geom_point() + facet_wrap(vars(dataset))
```
  These graphs are just the shapes described by their names.


### Problem 06    
   General physical activity, income, average hours worked per day. Basically, do museum attendees self-select for lower risk of mortality among known risk factors.


### Problem 10    
  f(b) = sum over i of (yi-B-xi). Take derivative of B with respect to B
-2 sum (yi-B-xi) = -2(sum(yi-xi) - n*B) = -2(ybar - xbar -nB) = 0
ybar-xbar = nB, (ybar-xbar)/n = B


### Problem 14    
  a. Yes. Intuitively, at least, denser urban areas seem to have lower school funding.
  b.
```{r}
a <- carData::Anscombe
cor(a$education, a$young)
```
  c. Slope is significant with p=0.026
```{r}
model <- lm(education ~ young, data=a)
summary(model)$coefficients
```
  d. From residuals, AK is a massive outlier. Updated model shows that the slope is not significantly different from 0.
```{r}
a<-a %>% filter(rownames(.)!="AK")
model <- lm(education ~ young, data=a)
summary(model)$coefficients
```

### Problem 15    
  a. 
```{r}
c <- fosdata::cern
c<-c %>% filter(platform=="Twitter") %>% head()
model<-lm(likes ~ shares, data=c)
plot(model)
```
  b. 
```{r}
l_model<-lm(log(likes) ~ shares, data=c)
plot(l_model)
```
  c. The log model looks a little better, but it is still pretty bad.   
  
  
### Problem 16    
  a.
```{r}
crit <- fosdata::crit_period
crit <- crit %>% filter(locale=="North America")
crit %>% ggplot(aes(x=aoa, y=gjt)) + geom_point()

```
  b. 
```{r}
model <- lm(gjt ~ aoa, data=crit)
model
crit %>% ggplot(aes(x=aoa, y=gjt)) + geom_point() + geom_smooth(method="lm")
```
   
  c. The slope represent the change is score based on the age someone learns a second language   
  d. No, there is no distinct bend in the graph.   
```{r}
plot(model)
```

### Problem 17    
  a.
```{r}
data(engel)
e<-engel
e %>%  ggplot(aes(x=income,y=foodexp)) + geom_point() + geom_smooth(method="lm")
```
   
  b.
```{r}
model<-lm(foodexp ~ income, data=e)
plot(model)
```
   
  c. Scale-location should be flat instead of trending upwards.

### Problem 20    
  P = 1.33*10^19, slope is significant, reject the null.
```{r}
pen <- palmerpenguins::penguins
pen<-pen %>% filter(species=="Gentoo")
model<-lm(body_mass_g ~ flipper_length_mm, data=pen)
summary(model)$coefficients

```

### Problem 21    
  a.   
```{r}
hot <- fosdata::hot_dogs
hot %>% ggplot(aes(x=calories,y=sodium, color=type)) + geom_point()
```
   
  b. Remove poultry the data looks to have 4 separate subgroups. Compare meat and beef instead. sodium = 3.613(calories) - 160.58   
```{r}
hot<-hot %>% filter(type!="Poultry")
model<-lm(sodium ~ calories, data=hot)
model
```
  c. 
```{r}
predict(model, newdata=data.frame(calories=140))

```
   
  d.   
```{r}
predict(model, newdata=data.frame(calories=140), interval="predict")

```

### Problem 22    
  a&b.
```{r}
bat<-Lahman::Batting
dat<-bat %>% group_by(yearID) %>% summarize(HR=sum(HR), R=sum(R))
dat %>% ggplot(aes(x=HR, y=R, color=yearID)) + geom_point()
```
  c.   
```{r}
dat<-dat %>% filter(yearID>1969)
dat %>% ggplot(aes(x=HR, y=R, color=yearID)) + geom_point()
```
   
  d. The slope predicts 2.54 runs for each homerun. The slope is significant with pval: 1.359*10^-23.
```{r}
model<-lm(R ~ HR, data=dat)
model
summary(model)$coefficients
```
   
  e. 
```{r}
dat %>% ggplot(aes(x=HR, y=R, color=yearID)) + geom_point() + geom_smooth(method="lm")
```
  f. ~19315, it would not be valid for predicting 1870 data. As we saw, there was a significant change in the relationship between HR and R around 1970.
```{r}
predict(model, newdata=data.frame(HR=4000))
```

### Problem 23    
  a.   
  b. Yes, there's a few. One Especially noticeable is one at ~51 stories that is close to 200 (height units?) taller than another 51 story building.   
  c. Not at a glance. They all seem to follow the same general trend.
```{r}
e <- Sleuth2::ex0728
e %>% ggplot(aes(x=Stories, y=Height, color=Year)) + geom_point() + geom_smooth(method="lm")
```
   
### Problem 24    
  a. The shape is roughly exponential, so the log of wine might help
```{r}
e <- Sleuth3::ex0823
e %>% ggplot(aes(x=Wine, y=Mortality)) + geom_point()
e %>% ggplot(aes(x=log(Wine), y=Mortality)) + geom_point()
```
   
  b. Yes, the data suggests there is a correlation between the two
```{r}
model <- lm(Mortality ~ log(Wine), data=e)
summary(model)$coefficients
cor(log(e$Wine), e$Mortality)

```
   
  c. No, because correlation does not imply causation.


### Problem 25    
```{r}
x <- runif(30, 0, 10)
epsilon <- rnorm(30)  # std normal, mean 0 sd 1
y <- x^2 + epsilon
plot(x,y)
mod <- lm(y ~ x)

# predict at x=5
pi <- predict(mod, data.frame(x=5), interval="predict")
rate<-replicate(10000000, {xnew <- 5;
ynew <- xnew^2 + rnorm(1);
# check the prediction
(pi[2] < ynew & pi[3] > ynew)})
mean(rate)

# predict at x=10
pi <- predict(mod, data.frame(x=10), interval="predict")
rate<-replicate(10000000, {xnew <- 10;
ynew <- xnew^2 + rnorm(1);
# check the prediction
(pi[2] < ynew & pi[3] > ynew)})
mean(rate)
```


### Problem 28    
  a&b.   
```{r}
x = seq(0,10,length.out = 21)

test<- replicate(10000, {epsilon <- rnorm(21, sd=3);
y = 1 + 2*x + epsilon;
model<-lm(y~x);
y[x==5.5] - predict(model, newdata=data.frame(x=5.5))})

sd(test)
```
   
  c.
```{r}
x = seq(0,10,length.out = 21)

test<- replicate(10000, {epsilon <- rnorm(21, sd=3);
y = 1 + 2*x + epsilon;
model<-lm(y~x);
y[x==10] - predict(model, newdata=data.frame(x=10))})

sd(test)
```
